---
title: "Analysis Corona"
author: "Meinari"
date: "3/10/2020"
output: 
  html_document:
    theme : cosmo
    highlight : tango
    toc: true
    toc_depth: 3
    toc_float: 
        collapsed: false
    number_sections: true
    fig_caption: yes
    df_print : paged 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(scipen = 9999)
rm(list=ls())
```

# Introduction
the data is about corona day-to-day data from 22-01-2020 until 4-03-2020, covering number of cases confirmed, number of deaths, number of heals accross the world. There are 3 research objectives of this analysis as followed : 
1. To compare corona spreading in China vs Non-China, along with its number of total recovered and death. 
2. To know heal rate and death rate of corona in each country. 
3. To see spreading of Corona across the world day-by-day. 


```{r}
#install library
library(readr) 
library(dplyr)
library(tidyr)
library(ggplot2)
library(glue)
library(plotly)
library(shiny)
library(shinydashboard)
library(flexdashboard)
library(tidyverse)
library(lubridate)
library(leaflet)
```


# 1. Read Data

```{r}
#read data corona and geojson

#data corona
corona <- read_csv("covid_19_clean_complete (21).csv")

# read data geojson
states <- geojsonio::geojson_read("custom.geo.json", what = "sp")
```

# 2. Exploratory Data Analysis
```{r}
# to see data structure
str(corona)

# to see if there is missing value of observation
colSums(is.na(corona))
```
```{r}
# change data structure for `Date`
corona$Date <- mdy(corona$Date)

# there are 3354 missing data of "province/states" so I decided to remove it from data frame
corona1 <- corona %>% 
  select(-`Province/State`) %>% 
  mutate(country = `Country/Region`,
         country = case_when(country =="Mainland China" ~ "China",
                             T ~ country),
         area = ifelse(country == "China", "China", "Not China"),
         area = as.factor(area))

# to input data to plygon (states) using leftjoin with variable `name` so need to match data corona. There are several countries that have different name. Therefore need to adjust first
corona1 <- corona1 %>% 
  mutate(country = case_when(country == "Czech Republic" ~ "Czech Rep.",
                             country == "Dominican Republic" ~ "Dominican Rep.",
                             country == "South Korea" ~ "Korea",
                             country == "United Arab Emirates" ~ "United Arab Emirate",
                             country == "UK" ~ "United Kingdom",
                             country == "US" ~ "United States",
                              T ~ country),
         name = as.factor(country),
         country = as.factor(country))

# groupby name and date

corona2 <- corona1 %>% 
  group_by(name) %>% 
  summarise(total_confirmed = sum(Confirmed),
            total_death = sum(Deaths),
            total_recovered = sum(Recovered)) %>% 
  ungroup() 


#left join data corona2 and states, joining by `name`
states@data <- left_join(x = states@data, y = corona2, by = "name")

# data for leaflet
bins <- c(1, 10, 100, 1000, 10000, Inf)
pal <- colorBin("YlOrRd", domain = states$total_confirmed, bins = bins)
```


# Data Analysis

## Objective number 1 
To compare corona spreading in China vs Non-China, along with its number of total recovered and death.

```{r}
corona11 <- corona1 %>% 
  group_by(Date) %>% 
  summarise(total_confirmed = sum(Confirmed),
            total_death = sum(Deaths),
            total_recovered = sum(Recovered)) %>% 
  ungroup() 

corona11_long <- pivot_longer(data = corona11, cols = -Date) %>% 
  mutate(case = name,
         text = glue(
    "Type Case: {case}
     Total Case = {value}"
  ))


plot1 <- ggplot(data = corona11_long, aes(x = Date, y = value)) +
  geom_line(aes(color = case), ) +
  geom_point(aes(color = case, text = text, show.legend= T)) +
  theme()+
  labs(title = "Corona Case 2020",
       x = "Date",
       y = "Total Case")

plot1
ggplotly(plot1, tooltip = "text")
```

## Objective Number 2 
To know heal rate and death rate of corona in each country.

```{r}
#Since the corona data is cummulative so we take the last date published (2020-03-04)

corona3 <- corona1 %>%
  group_by(country,Date) %>% 
  summarise(total_confirmed= sum(Confirmed),
            total_death = sum(Deaths),
            total_recovered = sum(Recovered)) %>% 
  ungroup %>% 
  filter(Date == "2020-03-04") %>% 
  mutate(death_ratio = total_death/total_confirmed,
         heal_ratio = total_recovered/total_confirmed) %>% 
  select(Date, heal_ratio, death_ratio)

corona3_long <- pivot_longer(data = corona3, cols = -Date) %>%
  mutate(ratio = name,
         text = glue(
    "Type Case: {ratio}
     Total Case = {value}"
  ))


plot2 <- ggplot(data = corona3_long, aes(x = Date, y = value, text=text)) +
  geom_col(aes(fill= ratio), show.legend = T)+
  coord_flip()+
  theme()+
  labs(title = "Death ratio vs Heal Ratio of COVID-19",
       subtitle = "Ratio by Country",
       x = "Date",
       y = "Case_ratio")

ggplotly(plot2, tooltip = "text")
```


## Objective Number 3 
To see spreading of Corona across the world day-by-day

```{r}
summary(corona1)
```


```{r}
m <- leaflet(states) %>% 
  addTiles("COVID-19 Confirmed Cases") %>% 
  setView(0,0,1.5) %>% 
  addPolygons(
  fillColor = ~pal(total_confirmed),
  weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7) %>% 
  addLegend(pal = pal, values = ~total_confirmed, opacity = 0.7, title = "Total Corona Confirmed",
  position = "bottomright")

m
```

```{r}
corona_ind <- corona1 %>%
  filter(`Country/Region` == "Indonesia",
         Date >= "2020-03-02") %>%
  group_by(Date) %>%
  summarise(total_confirmed = sum(Confirmed),
            total_death = sum(Deaths),
            total_recovered = sum(Recovered)) %>%
  ungroup()


coronaind_long <- pivot_longer(data = corona_ind, cols = -Date) %>%
  mutate(case = name,
         text = glue(
    "Type Case: {case}
     Total Case = {value}"
  ))


plot4 <- ggplot(data = coronaind_long, aes(x = Date, y = value, label = value)) +
  geom_line(aes(color = case), ) +
  geom_point(aes(color = case, text = text, show.legend= T)) +
  theme_bw()+
#   # theme(legend.position = c(0.1,0.8), legend.title = element_text(colour="black", size=14,
#   #                                     face="bold"), legend.text = element_text(size=20),legend.background = element_rect(fill="lightyellow", 
#   #                                 size=1, linetype="solid"))+
  geom_label(aes(fill = case), colour = "white", fontface = "bold")+
  labs(title = "COVID-19 Case in Indonesia",
       x = "Date",
       y = "Total Case")

plot4
ggplotly(plot4, tooltip = "text")

```


```{r}
corona11 <- corona1 %>% 
  group_by(Date) %>% 
  summarise(total_confirmed = sum(Confirmed),
            total_death = sum(Deaths),
            total_recovered = sum(Recovered)) %>% 
  ungroup() 

corona11_long <- pivot_longer(data = corona11, cols = -Date) %>% 
  mutate(case = name,
         text = glue(
    "Type Case: {case}
     Total Case = {value}"
  ))


plot1 <- ggplot(data = corona11_long, aes(x = Date, y = value)) +
  geom_line(aes(color = case), ) +
  geom_point(aes(color = case, text = text, show.legend= T)) +
  theme()+
  labs(title = "Corona Case 2020",
       x = "Date",
       y = "Total Case")

plot1
ggplotly(plot1, tooltip = "text")
```

```{r}
corona_china <- corona1 %>%
  filter(area == "China") %>% 
  group_by(Date) %>% 
  summarise(total_confirmed = sum(Confirmed),
            total_death = sum(Deaths),
            total_recovered = sum(Recovered)) %>% 
  ungroup() 

corona_china_long <- pivot_longer(data = corona_china, cols = -Date) %>% 
  mutate(case = name,
         text = glue(
    "Type Case: {case}
     Total Case = {value}"
  ))


plot_china <- ggplot(data = corona_china_long, aes(x = Date, y = value)) +
  geom_line(aes(color = case), ) +
  geom_point(aes(color = case, text = text, show.legend= T)) +
  theme_minimal()+
  theme(legend.position = c(0.1,0.8), legend.title = element_text(colour="black", size=14, 
                                      face="bold"), legend.background = element_rect(fill="lightblue", 
                                  size=0.5, linetype="solid"))+
  labs(title = "COVID-19 In China",
       x = "Date",
       y = "Total Case")

plot_china
ggplotly(plot_china, tooltip = "text")
```

```{r}
corona_out_china <- corona1 %>%
  filter(area == "Not China") %>% 
  group_by(Date) %>% 
  summarise(total_confirmed = sum(Confirmed),
            total_death = sum(Deaths),
            total_recovered = sum(Recovered)) %>% 
  ungroup() 

corona_out_china_long <- pivot_longer(data = corona_out_china, cols = -Date) %>% 
  mutate(case = name,
         text = glue(
    "Type Case: {case}
     Total Case = {value}"
  ))


plot_out_china <- ggplot(data = corona_out_china_long, aes(x = Date, y = value)) +
  geom_line(aes(color = case), ) +
  geom_point(aes(color = case, text = text, show.legend= T)) +
  theme_minimal()+
  theme(legend.position = c(0.1,0.8), legend.title = element_text(colour="black", size=14, 
                                      face="bold"), legend.background = element_rect(fill="lightblue", 
                                  size=0.5, linetype="solid"))+
  labs(title = "COVID-19 In Outside China",
       x = "Date",
       y = "Total Case")

plot_out_china
ggplotly(plot_out_china, tooltip = "text")

```

```{r}
corona_out_china <- corona1 %>%
  filter(area == "Not China") %>% 
  group_by(Date) %>% 
  summarise(total_confirmed = sum(Confirmed),
            total_death = sum(Deaths),
            total_recovered = sum(Recovered)) %>% 
  ungroup() 

corona_out_china_long <- pivot_longer(data = corona_out_china, cols = -Date) %>% 
  mutate(case = name,
         text = glue(
    "Type Case: {case}
     Total Case = {value}"
  ))


plot_out_china <- ggplot(data = corona_out_china_long, aes(x = Date, y = value)) +
  geom_line(aes(color = case), ) +
  geom_point(aes(color = case, text = text, show.legend= T)) +
  theme_bw()+
  labs(title = "COVID-19 In Outside China",
       x = "Date",
       y = "Total Case")

plot_out_china
ggplotly(plot_out_china, tooltip = "text")
```

